---
title: "analysis_final"
author: "Gabriel Saruhashi"
date: "4/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview of the data

Make a LIST of all variables you use –describe units, anything I should know.

```{r}
music = read.csv("brz_dictatorship.csv", as.is=TRUE)
dim(music)
str(music)

```
Create corpus for text mining
```{r}
library(tm)

jg <- paste(music$song_lyrics[music$class=="Jovem Guarda"], collapse = '')
protest <- paste(music$song_lyrics[music$class=="Protest"], collapse = '')
docs <- Corpus(VectorSource(c(jg, protest)))
#inspect(docs)


```

# Data Cleaning
describe the cleaning process you used on your data.  Talk about what issues you encountered.


```{r}
# Remove numbers
docs <- tm_map(docs, removeNumbers)
# Remove english common stopwords
docs <- tm_map(docs, removeWords, stopwords("portuguese"))
# Remove punctuations
docs <- tm_map(docs, removePunctuation)
# Eliminate extra white spaces
docs <- tm_map(docs, stripWhitespace)
# Remove your own stop word
# specify your stopwords as a character vector
docs <- tm_map(docs, removeWords, c("mim", "pra", "vai")) 
```

# Descriptive Plots & Summary Information
(Plots should be clearly labeled, well formatted,and display an aesthetic sense.)
Examine correlations between continuous variables

```{r}
library("ggplot2")
# 1. Open jpeg file
jpeg("valence_bp.jpeg", width = 350, height = 350)
boxplot(valence ~ class, data=music, col = 'yellow', main = "Valence by Class", ylab = "Valence")
#calculate means using the tapply function - could also use the by function
means <- tapply(music$valence, music$class, mean)
points(means, col = "red", pch = 19, cex = 1.2)
text(x=c(1:5), y=means+.2, labels = round(means,2))
dev.off()

jpeg("speechiness_bp.jpeg", width = 350, height = 350)
boxplot(speechiness ~ class, data=music, col = 'green', main = "Speechiness by Class", ylab = "Speechiness")
#calculate means using the tapply function - could also use the by function
means <- tapply(music$speechiness, music$class, mean)
points(means, col = "red", pch = 19, cex = 1.2)
text(x=c(1:5), y=means+.2, labels = round(means,2))
dev.off()

jpeg("energy.jpeg", width = 350, height = 350)
boxplot(energy ~ class, data=music, col = 'blue', main = "Energy by Class", ylab = "Energy")
#calculate means using the tapply function - could also use the by function
means <- tapply(music$energy, music$class, mean)
points(means, col = "red", pch = 19, cex = 1.2)
text(x=c(1:5), y=means+.2, labels = round(means,2))
dev.off()

jpeg("danceability.jpeg", width = 400, height = 400)
boxplot(danceability ~ class, data=music, col = 'pink', 
        main = "Danceability by Class", ylab = "Danceability")
#calculate means using the tapply function - could also use the by function
means <- tapply(music$danceability, music$class, mean)
points(means, col = "red", pch = 19, cex = 1.2)
text(x=c(1:5), y=means+.2, labels = round(means,2))
dev.off()
```

```{r}
print("Means by Music Class")
tapply(music$valence, music$class, mean)

print("SD by Class")
(sds <- tapply(music$valence, music$class, sd))

#Check ratio of largest to smallest sample sd
print("Ratio of Max/Min Sample SD")
round(max(sds)/min(sds),1)
```

```{r}
print("Means by Music Class")
tapply(music$speechiness, music$class, mean)

print("SD by Class")
(sds <- tapply(music$speechiness, music$class, sd))

#Check ratio of largest to smallest sample sd
print("Ratio of Max/Min Sample SD")
round(max(sds)/min(sds),1)
```
Kruskall-Wallis
```{r}
kruskal.test(music$speechiness ~ as.factor(music$class))
summary.aov(aov(music$speechiness ~ as.factor(music$class)))
```

## Lyric Analysis
Inspired by the analysis conducted by (http://www.sthda.com/english/wiki/text-mining-and-word-cloud-fundamentals-in-r-5-simple-steps-you-should-know)
```{r}
# Load
library("tm")
library("SnowballC")
library("wordcloud")
library("RColorBrewer")

# Remove numbers
docs <- tm_map(docs, removeNumbers)
# Remove english common stopwords
docs <- tm_map(docs, removeWords, stopwords("portuguese"))
# Remove punctuations
docs <- tm_map(docs, removePunctuation)
# Eliminate extra white spaces
docs <- tm_map(docs, stripWhitespace)

# Document matrix is a table containing the frequency of the words. Column names are words and row names are documents
dtm_jg <- TermDocumentMatrix(docs[1])
m <- as.matrix(dtm_jg)
v <- sort(rowSums(m),decreasing=TRUE)
d_jg <- data.frame(word = names(v),freq=v)
head(d_jg, 10)

dtm_protest <- TermDocumentMatrix(docs[2])
m <- as.matrix(dtm_protest)
v <- sort(rowSums(m),decreasing=TRUE)
d_protest <- data.frame(word = names(v),freq=v)
head(d_protest, 10)
```

Generate the worcloud for protest songs
```{r}
set.seed(1234)
wordcloud(words = d_protest$word, freq = d_protest$freq, min.freq = 15,
          max.words=200, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))


```
Generate wordclouds for Jovem Guarda
```{r}
wordcloud(words = d_jg$word, freq = d_jg$freq, min.freq = 15,
          max.words=200, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
```
Let's analyse now as barplots:
```{r}
par(mfrow=c(1,2))

barplot(d_jg[1:10,]$freq, las = 2, names.arg = d_jg[1:10,]$word,
        col ="orangered2", main ="Most frequent words for Jovem Guarda music",
        ylab = "Word frequencies")

barplot(d_protest[1:10,]$freq, las = 2, names.arg = d_protest[1:10,]$word,
        col ="lightblue", main ="Most frequent words for Protest music",
        ylab = "Word frequencies")
```
